Refractory anaphylaxis: Data from the European Anaphylaxis Registry

### 2018-09-18 analysis/paper/paper.Rmd
glucagon discussion add

Patients with refractory anaphylaxis more often reported concomitant proton pump inhibitors (PPI) and acetylsalicilic acid (ASA) use compared to patients with severe non-refractory anaphylaxis. Other (not specified) medications were also more frequent in refractory cases.

@@

@@

Definition  of refractory anaphylaxis is not universal. Also in this study, the confirmation of a refractory anaphylaxis based on the answers to our online questionnaire posed difficulties.
If a fatal reaction occured before the second dose of adrenaline was administered to the patient - it could not be diagnosed as refractory (as we defined at least two doses of minimum 300 µg adrenaline) although it in fact could have been refractory.

>>>>

Patients with refractory anaphylaxis more often reported concomitant proton pump inhibitors (PPI) and acetylsalicilic acid (ASA) use compared to patients with severe non-refractory anaphylaxis. Other (not specified) medications were also more frequent in refractory cases. In `r table(data3$rANA,data3$q_423_beta)[2,2]` cases of refractory anaphylaxis (`r (table(data3$rANA,data3$q_423_beta)[2,2]/sum(table(data3$rANA,data3$q_423_beta)[2,1:3])*100) %>% round(1)`%) patients reported receiving beta-blockers as a concomitant medication, but none of these patients recieved a glucagon infusion.

@@

Suprisingly, even though multiple anaphylaxis management guidelines recommend glucagon infusions in cases of concomitant beta-blocker therapy in anaphylactic patients [@Simons2011;@Lieberman;@Muraro2014], it has not been administered in any of the severe or refractory cases. Glucagon has been reported to sucessfully relive refractory anaphylaxis [@Zaloga1986;@Javeed1996] by directly activating the adenylyl cyclase and therefore bypassing the \beta-adrenergic receptor blockade [@Zaloga1986].

@@

Definition  of refractory anaphylaxis is not universal. Also in this study, the confirmation of a refractory anaphylaxis based on the answers to our online questionnaire posed difficulties.
If a fatal reaction occured before the second dose of adrenaline was administered to the patient - it could not be diagnosed as refractory (as we defined at least two doses of minimum 300 µg adrenaline for refractory anaphylaxis) although it in fact could have been refractory.
